---
title: "HW06"
author: "Chongdan Pan"
date: "2022/3/18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.align="center", error=FALSE, fig.width=10)
```

### Problem 6.1

Explore the data

```{r results="hide"}
library(doParallel)
library(doRNG)
library(tidyverse)
library(pomp)
library(ggplot2)
registerDoParallel()
registerDoRNG(2488820)

# Explore the data
read_csv(paste0("./Measles_Consett_1948.csv")) %>%
  select(week,reports=cases) -> meas
meas %>% as.data.frame() %>% head()
meas %>% ggplot(aes(x=week,y=reports)) + geom_line()+ geom_point()
```
sim_sir(c(Beta=7.5,mu_IR=0.5,rho=0.5,k=10,eta=0.03,N=38000))
sim_sir(c(Beta=28,mu_IR=0.3,rho=0.5,k=10,eta=0.03,N=38000))
sim_sir(c(Beta=28,mu_IR=0.3,rho=0.5,k=10,eta=0.03,N=38000))
sim_sir(c(Beta=28,mu_IR=0.3,rho=0.55,k=500,eta=0.03,N=38000))

Simulation of POMP model for following parameters

| Parameter | $\beta$ | $\mu_{IR}$ | $\rho$ | $k$  | $\eta$ | $N$   |
| --------- | ------- | ---------- | ------ | ---- | ------ | ----- |
| Figure 1  | 7.5     | 0.5        | 0.5    | 10   | 0.03   | 38000 |
| Figure 2  | 28      | 0.3        | 0.5    | 10   | 0.03   | 38000 |
| Figure 3  | 28      | 0.3        | 0.5    | 10   | 0.03   | 38000 |
| Figure 4  | 28      | 0.3        | 0.55   | 500  | 0.03   | 38000 |

Based on the previous result from slides, I increase the value of $\beta$ so that we have more reports. Then I slightly decrease the $\mu_IR$ so that the peak won't come so quickly. In the end, I changed the parameters for report gerneraton so that there is no too much volatility.


```{r results="hide"}
# Define the Pomp Model

sir_step <- Csnippet("
  double dN_SI = rbinom(S,1-exp(-Beta*I/N*dt));
  double dN_IR = rbinom(I,1-exp(-mu_IR*dt));
  S -= dN_SI;
  I += dN_SI - dN_IR;
  R += dN_IR;
  H += dN_IR;
  ")

sir_rinit <- Csnippet("
  S = nearbyint(eta*N);
  I = 1;
  R = nearbyint((1-eta)*N);
  H = 0;
  ")

sir_dmeas <- Csnippet("
  lik = dnbinom_mu(reports,k,rho*H,give_log);
  ")

sir_rmeas <- Csnippet("
  reports = rnbinom_mu(k,rho*H);
  ")

meas %>%
  pomp(
        times="week",t0=0,
        rprocess=euler(sir_step,delta.t=1/7),
        rinit=sir_rinit,
        rmeasure=sir_rmeas,
        dmeasure=sir_dmeas,
        accumvars="H",
        statenames=c("S","I","R","H"),
        paramnames=c("Beta","mu_IR","N","eta","rho","k")
       ) -> measSIR

### Analyze the result
sim_sir <- function (params){
  measSIR %>% simulate(
    params=params,
    nsim=20,format="data.frame",include.data=TRUE
  ) -> sims

  sims %>%
    ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
    geom_line()+
    guides(color=FALSE)
}

sim_sir(c(Beta=7.5,mu_IR=0.5,rho=0.5,k=10,eta=0.03,N=38000))
sim_sir(c(Beta=28,mu_IR=0.3,rho=0.5,k=10,eta=0.03,N=38000))
sim_sir(c(Beta=28,mu_IR=0.3,rho=0.5,k=10,eta=0.03,N=38000))
sim_sir(c(Beta=28,mu_IR=0.3,rho=0.55,k=500,eta=0.03,N=38000))
```

### 6.2 

Simulation of POMP model for following parameters.

| Parameter  | $\beta$ | $\mu_{IR}$ | $\rho$ | $k$  | $\eta$ | $N$   | $\mu_{EI}$ |
| ---------- | ------- | ---------- | ------ | ---- | ------ | ----- | ---------- |
| Figure 1   | 28      | 0.3        | 0.55    | 500  | 0.03   | 38000 | 0.5        |
| Figure 2   | 40      | 0.35        | 0.5    | 500  | 0.03   | 38000 | 0.9    |

Firstly, I starts with the parameters from previous question and setting the transition rate from exposed to infectious $\mu_{EI}$ to be 0.5. The result is not good enough because the peak is low and coming too late. What's more, there are too much noise after the peak. Therefore, I tuned the parameters of $\mu_{EI},\mu_{IR},\rho$ to make it better.

```{r results="hide"}
sir_step <- Csnippet("
  double dN_SE = rbinom(S,1-exp(-Beta*I/N*dt));
  double dN_EI = rbinom(E,1-exp(-mu_EI*dt));
  double dN_IR = rbinom(I,1-exp(-mu_IR*dt));
  S -= dN_SE;
  E += dN_SE - dN_EI;
  I += dN_EI - dN_IR;
  R += dN_IR;
  H += dN_IR;
  ")

sir_rinit <- Csnippet("
  S = nearbyint(eta*N);
  E = 0;
  I = 1;
  R = nearbyint((1-eta)*N);
  H = 0;
  ")

sir_dmeas <- Csnippet("
  lik = dnbinom_mu(reports,k,rho*H,give_log);
  ")

sir_rmeas <- Csnippet("
  reports = rnbinom_mu(k,rho*H);
  ")

meas %>%
  pomp(
        times="week",t0=0,
        rprocess=euler(sir_step,delta.t=1/7),
        rinit=sir_rinit,
        rmeasure=sir_rmeas,
        dmeasure=sir_dmeas,
        accumvars="H",
        statenames=c("S","E","I","R","H"),
        paramnames=c("Beta","mu_IR","N", "eta", "rho", "k", "mu_EI")
       ) -> measSIR

### Analyze the result
sim_seir <- function (params){
  measSIR %>% simulate(
    params=params,
    nsim=20,format="data.frame",include.data=TRUE
  ) -> sims

  sims %>%
    ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
    geom_line()+
    guides(color=FALSE)
}


sim_seir(c(Beta=28,mu_IR=0.3,rho=0.55,k=500,eta=0.03,N=38000,mu_EI=0.5))
sim_seir(c(Beta=40,mu_IR=0.35,rho=0.5,k=500,eta=0.03,N=38000,mu_EI=0.9))
```

### Refernce

- [Slides about Chapter 12](https://ionides.github.io/531w22/12/index.html)

- [Compartmental models in epidemiology](https://en.wikipedia.org/wiki/Compartmental_models_in_epidemiology)

- [Previous Solution](https://ionides.github.io/531w21/hw06/sol06.html)