---
title: "HW09"
author: "Chongdan Pan"
date: '2022-04-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Question 1

$$
\gamma(h)=\text{Cov}(X_{n+h},X_n)=\text{Cov}(\sum_{i=-\infty}^\infty\psi_i\epsilon_{n-i},\sum_{j=-\infty}^\infty\psi_j\epsilon_{n+h-j})\\
\gamma(h)=\text{Cov}(\sum_{i=-\infty}^\infty\psi_i\epsilon_{n-i},\sum_{i=-\infty}^\infty\psi_{i+h}\epsilon_{n-i})=\sum_{i=-\infty}^\infty\psi_i\psi_{i+h}\sigma^2
$$

## Question 2

(a)
$$
(1-\alpha_1B)X_n=(1+\theta_1B+\theta_2B^2)\epsilon_n\\
X_n=\frac{1+\theta_1B+\theta_2B^2}{1-\alpha_1B}\epsilon_n=\sum_{i=0}^\infty\alpha_1^iB^i(1+\theta_1B+\theta_2B^2)\epsilon_n=\sum_{i=0}^\infty\alpha_1^i(1+\theta_1B+\theta_2B^2)\epsilon_{n-i}\\
X_n=\sum_{i=0}^\infty\alpha_1^i(\epsilon_{n-i}+\theta_1\epsilon_{n-i-1}+\theta_2\epsilon_{n-i-2})=\epsilon_n+(\theta_1+\alpha_1)\epsilon_{n-1}+\sum_{i=0}^\infty\alpha_1^i(\theta_2+\alpha_1\theta_1+\alpha_1^2)\epsilon_{n-2-i}
$$

(b)
Based on the result from question 1, we have

$$
\gamma(h)=\sum_{i=-\infty}^\infty\psi_i\psi_{i+h}\sigma^2=\sum_{i=0}^\infty\psi_i\psi_{i+h}\sigma^2\\
\text{When } h=0\quad\gamma(h)=[1+(\alpha+\theta_1)^2+\sum_{i=0}^\infty\alpha_1^{2i}(\alpha_1^2+\alpha_1\theta_1+\theta_2)^2]\sigma^2 \\
\text{When } h=\pm1\quad \gamma(h)=[(\alpha_1+\theta_1)(1+\alpha_1^2+\alpha_1\theta_1+\theta_2)+\sum_{i=0}^\infty\alpha_1^{2i+1}(\alpha_1^2+\alpha_1\theta_1+\theta_2)^2]\sigma^2 \\
\text{When } h \geq 2 \text{ or } h\leq-2\quad \gamma(h)=\sum_{i=0}^\infty(\theta_2+\alpha_1\theta_1+\alpha_1^2)[1+\theta_1\alpha_1+(1+\theta_2)\alpha_1^2+\alpha_1^3\theta_1+\alpha_1^4]\alpha_1^{|h|-2+i}\sigma^2\\
$$

## Question 3

```{r}
X = read.csv("../NYA-2015-2017.csv",header=TRUE)
library(tseries)
NYSE_lret = diff(log(X$AdjClose))
NYSE_lret.ts <- ts(data=NYSE_lret,start=c(2015,1),frequency=252,names=c('logret'))
```

(a)

```{r}
plot(NYSE_lret.ts)
plot(abs(NYSE_lret.ts))
```

Based on the plot,  it looks like we can use a ARCH model to fit the absolute return, which is volatility because there is chance that the volatility at time $t$ may depend on it at $t-1$.

(b)

```{r}
model <- garch(NYSE_lret.ts, order=c(1,1))
summary(model)
```

For standard error, We have $6.23\times10^{-7}$ for $\alpha_0$, $2.04\times10^{-2}$fro $\alpha_1$, $2.73\times10^{-2}$ for $\beta_1$

(c)

```{r}
csd <- fitted.values(model)
plot(csd[,1])
title(main="Estimated Condition Volatilities")
resid <- residuals(model)
plot(resid)
title(main="Estimated Residuals")
```

(d)

```{r}
sd(na.omit(resid))
mean(na.omit(resid))
plot(model)
```

The mean and standard deviation of the residual is close to 0 and 1 respectively, implying that our code is working/ However, from the Q-Q plot of the residuals, it turns out we have heavy tails, which conform to the result from the Jarque Bera test. Since the P-value of the Box-Ljung test is large, we don't have significant evidence showing that the residuals are auto-correlated, which also conforms to the ACF plot. The value of $\alpha_0$ is quite small, implying that the lower bound of the volatility is small. In summary, our model does capture the time-series relation within the volatility, but we still need some way to deal with the heavy tail residuals.


(e)

Since the return's mean is extremely close to 0, we will assume $\mu_n=0$ in this case.

```{r}
sigma = (coef(model)[1] + coef(model)[3] * csd[754] ^ 2 + coef(model)[2] * NYSE_lret.ts[754] ^ 2) ^ 0.5
var = exp(-qnorm(0.005) * sigma) - 1
var

```

Therefore we have relative VaR to be 0.0118

(f)

```{r warning}
q <- (1:length(NYSE_lret.ts)) / (length(NYSE_lret.ts) + 1)
nu <- 4
qqplot(resid, qt(q, nu))
qqline(resid, qt(q, nu))
```
I'm using a t-distribution with degree of freedom to be 4 for fitting. Although we have a thin right tail, the left tail is well-fit, which is critical to the calculation of VaR.

```{r}
lambda = sqrt(sigma^2*(nu-2)/nu)
var = exp(-qt(0.005, nu) * lambda) - 1
var
```

Therefore, we have relative VaR based on T-distribution to be 0.0149

(g)

```{r}
day = which.max(csd)
day
sigma = (coef(model)[1] + coef(model)[3] * csd[day-1] ^ 2 + coef(model)[2] * NYSE_lret.ts[day-1] ^ 2) ^ 0.5
var = -qnorm(0.005) * sigma
var = exp(var) - 1
var
```

The index we have max volatility is 165 Therefore we have relative VaR to be 0.056 from normal distributed residual.

```{r}
lambda = sqrt(sigma^2*(nu-2)/nu)
var = exp(-qt(0.005, nu) * lambda) - 1
var
```

Therefore, we have relative VaR based on T-distribution to be 0.071

