---
title: "Stat 531 Homework 8"
author: "Abby Loe, Chongdan Pan, Hanjun Li"
date: "March 2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "50%", cache = TRUE)
library(ggplot2)
library(plyr)
library(dplyr)
library(broom)
library(car)
library(pomp)
library(tidyverse)
```

# Question 1

(A) is correct because the measurement time needs to match the data, while the markov process can have as small of time steps as we want.

# Question 2

(A) is correct because W's type was not declared. Also, by process of elimination, none of the other possible reasons make any type of remote sense.


# Question 3
(B) is correct because we have an object that should be passed to the pomp object, but it doesn't have the parameter names that it needs.


# Question 4

(A) is correct; rnorm is an R command, and it has been used in a C-snippet. The C compiler cannot recognize the R command, and instead has started to treat it like a parameter passed to a C function.


# Question 5

i) and ii) are not true; we could choose $h(V_n)=3$, which would not have the Markov property, and we could choose $h(V_n)=I$ which would make $W_n$ a Markov process. This means that for some choices, $W_n$ is a Markov process.

For a POMP model, $X_n$ is the latent variable, and is a Markov process because it's next state only depends on the previous state. Therefore iv) is true. This makes our final answer (B).


# Question 6



# Question 7



# Question 8
```{r}

d <- dacca(deltaI=0.08)
d%>%
  pfilter(Np=1000) -> pf
pf %>% logLik() %>% logmeanexp(se=FALSE) -> L_pf
round(L_pf)
```



# Question 9

# Question 10

# Question 11


# Question 12

# Question 13



# Question 14



# Question 15



# Question 16