---
title: "HW02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Problem 1

#### (a)

Based on the definition of $\text{P}[R_t\leq-\tilde{\text{VaR}}]=q$, and
$$
F_{\text{dexp}}(x)=\begin{cases}
\frac{1}{2}e^{\lambda(r-\mu)} & r<\mu\\
1-\frac{1}{2}e^{\lambda(\mu-r)} & r\geq\mu
\end{cases}
$$


<u>We have</u> 
$$
\tilde{\text{VaR}}_q=-r=-\frac{1}{\lambda}\ln2q-\mu
$$

#### (b)

We can change the the definition of $ES_q$ to be based on the return where $ES_q=-\frac{1}{q}\int_{x\leq-\text{VaR}_q}xf(x)\mathrm d x$,

Since $-\tilde{\text{VaR}}_q\leq\mu$, we have
$$
ES_q=-\frac{\lambda}{2q}\int_{-\infty}^\tilde{-\text{VaR}_q}xe^{\lambda(x-\mu)}\mathrm dx=-\frac{e^{-\lambda\mu}\lambda}{2q}\int_{-\infty}^\tilde{-\text{VaR}_q} xe^{\lambda x}\mathrm dx
$$
​	For the integral, we have
$$
\int_{-\infty}^\tilde{-\text{VaR}_q} xe^{\lambda x}\mathrm dx=-\frac{1+\lambda\tilde{\text{VaR}}_q}{\lambda^2}e^{-\lambda \tilde{\text{VaR}}_q}=\frac{\lambda\mu+\ln2q-1}{\lambda^2}e^{\ln 2q+\lambda\mu}
$$
​	<u>Plugging the integral, we have</u> 
$$
ES_q=\frac{1-ln2q}{\lambda}-\mu
$$

#### (c)

First we calculate the mean and deviation of the sample, and use them to estimate mu and lambda of the distribution

```{r}
df <- read.csv("../Nasdaq.csv", header=TRUE)
Return = df$Adj.Close[2:nrow(df)] / df$Adj.Close[1:nrow(df)-1] - 1
mu = mean(Return) 
std = sd(Return)
lambda = (2 ** 0.5) / std
```

Therefore we get $\mu=0.001260808$ and $\lambda=90.83559$

Since $q=0.01\leq\frac{1}{2}e^{-2\mu\lambda}\approx0.4$, we have

```{r}
if(0.5 * exp(-2 * mu * lambda)>0.1){
    VaR = 1e7 * (-log(0.02) / lambda - mu)
    ES = 1e7 * ((1-log(0.02)) / lambda - mu)
}
```

- $\text{VaR}_q\approx418062.7$
- $ES_q\approx528151.7$

<u>The Value at Risk means at probability less than $1\%$ of suffering a loss greater than 418062.7</u>

#### (d)

For normal distribution, we have

```{r}
VaR = 1e7 * -qnorm(p=0.01, mean=mean(Return), sd = sd(Return))
ES = 1e7 * (mu + std * dnorm(qnorm(0.01))/ 0.01)
```

- $\text{VaR}_q\approx353015.7$
- $ES_q\approx427553.6$

<u>I think result from double exponential distribution is more accurate since we have a high kurtosis result from last homework, meaning the return have a heavier tail than normal distribution and a higher risk.</u>

## 2. Problem 2

#### (a)

Based on the definition of Value of Risk, we have 
$$
P[R_t<-\tilde{\text{VaR}}_t]=P[X>\tilde{\text{VaR}}_q]=1-P[X\leq\tilde{\text{VaR}}]=q
$$
Therefore, the $\tilde{\text{VaR}}_q$ is the $1-q$ quantile of the loss

Since $P(X\leq\mu)=0.9$, then $1-(1+\frac{\xi(x-\mu)}{\sigma})^{-1/\xi}=(1-q-0.9)/(1-0.9)=1-10q$

<u>Therefore</u> $\tilde{\text{VaR}}_q=\mu-\frac{\sigma}{\xi}+\frac{\sigma}{\xi(10q)^\xi}=\mu-\frac{\sigma}{\xi}+\frac{\sigma}{\xi0.1^\xi}$

#### (b)

Based on the definition of shortfall,
$$
\Theta_{X_q}=P[X\leq x|X\geq\tilde{\text{Var}_q}]
$$
When $x<\tilde{\text{Var}_q}$, $\Theta_{X_q}=0$ 

otherwise
$$
\Theta_{X_q}(x)=\frac{\Theta_X(x)-\Theta_X(\tilde{\text{Var}_q})}{1-\Theta_X(\tilde{\text{Var}_q})}=\frac{\Theta_X(x)-(1-q)}{q}=1-\frac{1}{q}(1-\Theta_X(x))
$$
Since we know $\tilde{\text{VaR}}_q>\mu$, we can plug in the distribution of $\Theta_X(x)=0.9+0.1[1-(1+\frac{\xi(x-\mu)}{\sigma})^{-1/\xi}]$
$$
\Theta_{X_q}(x)=1-\frac{1}{10q}(1+\frac{\xi(x-\mu)}{\sigma})^{-1/\xi}=1-[(10q)^\xi+\frac{(10q)^\xi\xi(x-\mu)}{\sigma}]^{-1/\xi}
$$
We can reorganize the CDF to update $\mu'$ and $\sigma'$
$$
\Theta_{X_q}(x)=1-[\frac{\sigma+\xi(x-\mu)}{\sigma/(10q)^\xi}]^{-1/\xi}=1-(1+\frac{\xi[x-\mu+\sigma/\xi-\sigma/(\xi(10q)^\xi)]}{\sigma/(10q)^\xi})^{-1/\xi}
$$
<u>Therefore we have the new parameters</u> $\xi'=\xi,\sigma'=\frac{\sigma}{(10q)^\xi},\mu'=\mu-\frac{\sigma}{\xi}+\frac{\sigma}{\xi(10q)^\xi}=\tilde{\text{VaR}}_q$

#### (c)

Based on the expectation of generalized Pareto distribution
$$
E(X)=\int_{-\infty}^\infty x\theta_q(x)\mathrm dx
$$
Since $f(x)=0\quad\forall x\leq\mu$, then
$$
E(X)=\int_{\mu}^\infty x\theta_q(x)\mathrm dx=\int_{\tilde{\text{Var}_q}}^{\infty}x\theta_q(x)\mathrm dx
$$
Based on the definition of Expected shortfall
$$
ES_q=\frac{1}{q}\int_{\tilde{\text{Var}_q}}^{\infty}xf(x)\mathrm dx=\int_{\tilde{\text{Var}_q}}^{\infty}x\theta_q(x)\mathrm dx=E(X)
$$

<u>Therefore, we have</u>
$$
ES_q=\mu-\frac{\sigma}{\xi}+\frac{\sigma}{\xi(10q)^\xi}+\frac{\sigma}{(1-\xi)(10q)^\xi}
$$



